\chapter{LITERATURE REVIEW}

\section{Evolution of Terrain Representation}
For most of the twentieth century, topographic data came from field surveys and photogrammetric mapping \cite{usery2019topographic}. These methods could not produce the dense elevation grids needed for automated skyline computation.

The Shuttle Radar Topography Mission (SRTM) in 2000 delivered the first near-global elevation dataset \cite{Franks2023}. The ASTER Global DEM followed in 2009 \cite{pan2020camera}. Both datasets were useful but unreliable in steep terrain---radar shadows and layover artifacts introduced large errors around cliffs and deep valleys \cite{pan2020camera, Franks2023}, leading to inaccurate peak heights \cite{pan2020camera}.

The Copernicus GLO-30 model, released by the European Space Agency (ESA) in 2020 \cite{Franks2023, Trevisani2023}, resolved many of these problems. Independent evaluations confirm its superior vertical accuracy in steep mountain terrain \cite{Franks2023}. At 30-meter resolution, GLO-30 captures the distinctive shapes of Himalayan peaks while remaining small enough to store and process on portable devices \cite{pan2020camera, Steger2022}.

\section{Developments in Skyline Segmentation}
Early skyline extraction methods in the 2000s relied on edge-detection filters that identified sharp changes in image intensity \cite{yang2007robust}. These filters performed poorly when haze, clouds, or shadows obscured the sky--terrain boundary \cite{pan2020camera}.

The introduction of fully convolutional networks in 2015 marked a turning point for image segmentation \cite{long2015fully}. Deep models could learn to distinguish sky from terrain even under poor lighting. However, the earliest networks were computationally heavy. Lighter architectures such as MobileUNet addressed this by using depthwise separable convolutions, which cut computation while preserving pixel-level accuracy \cite{howard2019mobilenetv3}. More recent frameworks like YUNet adapt object detection to rapid sky segmentation and perform well under difficult weather conditions \cite{Yang2025}.

A recurring obstacle has been the scarcity of labeled training data. The GeoPose3K dataset, introduced by Brejcha and Cadik in 2017, provides over 3,000 alpine mountain images with precise camera pose labels \cite{Brejcha2017}. Tomesek et al.\ later explored cross-modal rendering to generate additional synthetic training examples \cite{tomesek2022cross}.

\section{Progress in Silhouette Matching}
Once a skyline is extracted, it must be matched against the reference database to determine the camera's position and heading \cite{pan2020camera}.

Early matching approaches compared profiles pixel by pixel, which was both slow and fragile---different camera lenses and viewpoints easily broke the alignment \cite{tonge2014parsing}. The method also assumed the camera's heading was already known.

Rotational sliding-window search removed the heading requirement by testing the query against all possible angular offsets of each reference profile \cite{pan2020camera}. To make this tractable, researchers adopted two-stage pipelines. In 2013, Rakthanmanon et al.\ showed that Dynamic Time Warping (DTW) could be applied to fast subsequence matching \cite{Rakthanmanon2013}. DTW is tolerant of the stretching and compression caused by different focal lengths \cite{Rakthanmanon2013}.

Full DTW search across a large database remains expensive \cite{Rakthanmanon2013}. Mueen et al.\ addressed this with a distance-profile algorithm that quickly eliminates poor candidates, leaving only a handful for DTW alignment \cite{mueen2024mass}. This two-stage combination---fast filtering followed by elastic matching---makes real-time localization feasible on mobile hardware.

Baatz et al.\ introduced the first complete large-scale geolocation pipeline for mountains in 2012 \cite{baatz2012large}. The system used skylines from elevation models to find unknown camera headings across the whole Alps. Saurer et al.\ later released a journal extension that scaled this matching approach further \cite{saurer2016image}. The CrossLocate project later localized 39\% of Alpine images within a one-kilometer radius \cite{tomesek2022cross}. The researchers tested all three systems exclusively in the European Alps.

\section{Validation with Geotagged Imagery}
Any alternative positioning system needs ground truth with reliable coordinates. Early studies used small, custom photo sets that lacked diversity in lighting and precise GPS tags.

The Google Street View Trekker campaign along the Everest trail \cite{google2015streetview} changed this. Captured in 2014, it contains over 45,000 high-resolution panoramas with accurate location and orientation metadata \cite{google2015streetview}. The Khumbu trails covered by this dataset provide a large, georeferenced collection of real mountain views suitable for systematic testing.

DEM-based skyline matching has been proven effective in mountainous 
terrain \cite{baboud2011automatic, baatz2012large, saurer2016image, 
pan2020camera, tomesek2022cross}. However, all existing systems target 
European or Chinese terrain. No system has been built or tested in the 
Himalayas, where extreme elevation, absent training data, and GPS 
failure in deep valleys present a distinct and unsolved problem. 
Table~\ref{tab:litreview} summarizes existing approaches.

\begin{table}[h]
\centering
\caption{Summary of existing skyline geolocation approaches}
\label{tab:litreview}
\begin{tabular}{llll}
\toprule
Authors & Year & Terrain & Matching Method \\
\midrule
Baboud et al. \cite{baboud2011automatic} & 2011 & Alps & Edge cross-correlation \\
Baatz et al. \cite{baatz2012large} & 2012 & Alps (Switzerland) & Bag-of-curves \\
Saurer et al. \cite{saurer2016image} & 2016 & Alps & Contours + DTW \\
Pan et al. \cite{pan2020camera} & 2020 & China (hilly) & Euclidean + heatmap \\
Tomešek et al. \cite{tomesek2022cross} & 2022 & Alps & Synthetic rendering \\
\midrule
\textbf{This work} & \textbf{2026} & \textbf{Khumbu, Nepal} & \textbf{FFT + DTW} \\
\bottomrule
\end{tabular}
\end{table}

\section{Research Gap}
Existing systems work well in the Alps or in Chinese hills \cite{baboud2011automatic, baatz2012large, saurer2016image, tomesek2022cross, pan2020camera}. No researchers have built or tested these pipelines in Himalayan terrain. The Himalayas present unique challenges including extreme elevation changes and an absence of labeled training data. This project adapts the established pipeline to Khumbu terrain using synthetic training data generated from the DEM.
